Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-08-30 mu/article/prb-sent1/manuscript.tex
First manuscript sent to PRB

%\documentclass[prb,11pt,tightenlines,twocolumn,aps]{revtex4-1}
\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
%\documentclass[preprint,prb,showpacs]{revtex4}
\usepackage{amsmath}    % need for subequations 
\usepackage{graphicx}   % need for figures 
\usepackage{verbatim}   % useful for program listings 
\usepackage{color}      % use if color is used in text 
\usepackage{subfigure}  % use for side-by-side figures 
\usepackage[pdffitwindow,colorlinks,citecolor={red},linkcolor={blue}]{hyperref}
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage{soul}
\def\tama{10cm}
\raggedbottom           % don't add extra vertical space

@@

\title{Pure Spin Current Injection in Hydrogenated Graphene Structures}
\author{Reinaldo Zapata-Pe\~na}
\affiliation{Centro de Investigaciones en \'Optica, Le\'on,
Guanajuato 37150, M\'exico}
\author{Bernardo S. Mendoza}
\affiliation{Centro de Investigaciones en \'Optica, Le\'on,
Guanajuato 37150, M\'exico}
\author{Anatoli I. Shkrebtii}
\affiliation{University of Ontario, Institute of Technology,
Oshawa, ON, L1H 7L7, Canada}

@@

@@

Spintronics is an emerging research field of electronics in which the
manipulation and transport of the electron spin in a solid state materials  is
central, adding a new degree of freedom to conventional charge
manipulation.\cite{wolfSC04,fabianAPS07} At present, there is an increasing
interest in attaining the same level of control over the transport of spin at
micro- or nano-scales, as it has been done for the flow of charge in typical
3D-bulk based electronic devices.\cite{awschalomNP2007} Several semiconductor
spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC). One of
the difficulties in creating measurable  spin current and
development of PSC based semiconductor devices is the fact that the spin
relaxation time in conventional semiconducting materials is too short to enable the
spin transport, thus resulting in a non-observable spin
current.\cite{murakamiSc03} For PSC there is no net motion of charge; spin-up
electrons move in a given direction, while spin-down electrons travel in the
opposite one. This effect can be due to one-photon absorption of linearly
polarized light by a semiconductor, with filled valence bands and empty
conduction bands, illuminated by light with photon energy larger than the
energy gap. This phenomenon can be due to spin injection,\cite{malPRB03} Hall
Effects,\cite{sinovaPRB04} interference of two optical beams,\cite{bhatPRL00,
najmaiePRB03} or one photon absorption of linearly polarized
light\cite{bhatPRL05}. The last effect has been observed in gallium arsenide
(GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

@@

In the top panel of Fig. \ref{fig:up-vsz-w1}, we plot
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ {\it vs.}
0.080\,eV$\leq\hbar\omega\leq$0.096\,eV (similar energy range for the Up
structure shown in the left panel of Fig.~\ref{fig:vab-str-comp}) and
$0^\circ\leq\alpha\leq 180^\circ$. We see a broad peak that reaches the maximum
of $\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 739.7$\,Km/s at
$\alpha=35^{\circ}$ and $\hbar\omega= 0.084$\,eV.  The variation of
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ as a function of $\alpha$,
which comes from the interplay of the $\boldsymbol{\mu}$ tensor components as
multiplied by the  trigonometric functions of Eq.~\eqref{eq:vab-aw}, gives a
sizable set of values between 739.7\,Km/s and 165.4\,Km/s, for
0.084\,eV$\leq\hbar\omega\leq$0.090\,eV. In the bottom panel, we show
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ (left scale,
black solid line) {\it vs.} $\alpha$, at $\hbar\omega= 0.084$\,eV, thus following the ridge in the 3D
plot of the top panel. Also, we plot the corresponding velocity angle
$\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)$ (right scale, red
short-dashed line), where it
is very interesting to see that $\gamma_{\sigma^z}(\omega,\alpha)$ is centered
at 64.55$^\circ$ with a rather small deviation of only $\pm 0.03^\circ$, for
the whole range of $\alpha$. This result means that for $\hbar\omega=0.084$ eV
and for all values of $\alpha$, the electrons, with the chosen spin pointing
along $z$, will move at the angle of $\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)
\sim 64.5^{\circ}$ with respect to the $x$ direction, with the range of  high
speeds $\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ shown in the figure.
Also, from Eq. \eqref{eq:gamma-par} we find that
$\gamma^\parallel_{\sigma^\mathrm{z}} (\omega,\alpha)=\alpha = 64.56^\circ$,
with $\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha) = 631.1$\,Km/s 
(as indicated by the green dot-dashed arrow), and that from Eq. \eqref{eq:gamma-perp},
$\gamma^\perp_{\sigma^\mathrm{z}}(\omega,\alpha)=\alpha-90^\circ=64.50^\circ$,
gives $\alpha=154.50^\circ$, with
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 191.5$ Km/s 
(as indicated by the orange long-dashed 
arrow). 
Thus, at $\hbar\omega=0.084$ eV, an incident field, polarized at
$\alpha \sim 65.5^\circ$ or $\sim 154.5^\circ$, injects electrons with their
spin polarized along $z$, which move parallel or perpendicular to the incident
electric field,  with a speed of 631.14\,Km/s or 191.5\,Km/s, respectively.

Now, we analyze the results for the second energy range of the Up structure
shown in Fig.~\ref{fig:vab-str-comp}. In the top panel
of Fig. \ref{fig:up-vsz-w2}, we plot $\mathcal{V}_{\sigma^{\mathrm{z}}}
(\omega,\alpha)$ {\it vs.} 1.950\,eV$\leq\hbar\omega\leq$1.960\,eV and
$0^\circ\leq\alpha\leq 180^\circ$. We see two broad peaks that maximize at
$\alpha=35^{\circ}$ and $\hbar\omega= 1.954$\,eV, with a value of
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 193.5$\,Km/s, and at
$\alpha=35^{\circ}$ and $\hbar\omega= 1.957$\,eV, with a value of
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 170.6$\,Km/s. We only
analyze the highest  maximum in the bottom panel, where we  show
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ (left scale,
black solid line) {\it vs.} $\alpha$, at $\hbar\omega= 1.954$\,eV, thus following the highest ridge
shown in the 3D plot of the top panel. Also, we plot the corresponding velocity
angle $\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)$ (right scale, red short-dashed  line),
where in this case we see that the values of $\gamma_{\sigma^z}(\omega,\alpha)$
have more dispersion, as a function of $\alpha$, than for the lower energy
range shown in the bottom panel of Fig.~\ref{fig:up-vsz-w1}. However,
$\gamma_{\sigma^z}(\omega,\alpha)\sim 77.8^\circ$ is nearly constant from
$\alpha=0^\circ$ up to $\alpha\sim 85^\circ$. In this case, we find that
$\gamma^\parallel_{\sigma^\mathrm{z}}(\omega,\alpha)=\alpha=78.0^\circ$, with
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 115.0$\,Km/s 
(as indicated by the green dot-dashed arrow), and that from Eq. \eqref{eq:gamma-perp},
$\gamma^\perp_{\sigma^\mathrm{z}}(\omega,\alpha)=\alpha-90^\circ=167.8^\circ$,
gives $\alpha=77.8^\circ$, with
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 65.6$\,Km/s 
(as indicated by the orange long-dashed  arrow). 
Thus, through the correct choice of $\hbar\omega$ and $\alpha$ we could
inject electrons, in this case with their spin polarized along $z$, which move
parallel or perpendicular to the incident electric field, with sizable speeds.

@@

We proceed to analyze the Alt structure, just as we did with the Up structure,
but in this case, we only choose the lower energy range shown in the left
central panel of Fig.~\ref{fig:vab-str-comp}. In the top panel of Fig. 
\ref{fig:alt-vsz}, we plot $\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$
{\it vs.} 0.715\,eV$\leq\hbar\omega\leq$0.725\,eV and $0^\circ\leq\alpha\leq
180^\circ$. We see a broad peak that maximizes at $\alpha=150^{\circ}$ and
$\hbar\omega= 0.720$\,eV, with a value of
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 644.9$\,Km/s. In the bottom
panel, we  show $\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ (left scale, black solid line) {\it vs.}
$\alpha$, at $\hbar\omega= 0.720$\,eV, thus following
the highest ridge shown in the 3D plot of the top panel. Also, we plot the
corresponding velocity angle $\gamma_{\sigma^\mathrm{z}} (\omega,\alpha)$
(right scale, red short-dashed  line), where now we see that
$\gamma_{\sigma^z}(\omega,\alpha)$ is centered at $109.2^{\circ}$ having
variations of $\pm 1.0^{\circ}$ for $0^{\circ} \leq
\alpha \leq 180^{\circ}$. In this case, we find that
$\gamma^\parallel_{\sigma^\mathrm{z}} (\omega,\alpha) =
\alpha = 108.8^\circ$, with $\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)
= 450.05$\,Km/s (as indicated by the  green dot-dashed arrow), 
and that from Eq. \eqref{eq:gamma-perp},
$\gamma^\perp_{\sigma^\mathrm{z}}(\omega,\alpha)=\alpha-90^\circ=110.0^\circ$,
gives $\alpha=20.0^\circ$, with $\mathcal{V}_{\sigma^{\mathrm{z}}}
(\omega,\alpha) = 60.84$\,Km/s 
(as indicated by the orange long-dashed  arrow). Thus, as for the Up structure,
we could inject electrons with a fixed spin, which moves parallel or
perpendicular to the incident electric field.
% Again, for the cases in which the spin polarization is parallel to the
% surface the maximum values are
% $\mathcal{V}_{\sigma^{\mathrm{x}}}(\omega,\alpha) = 297.8$ Km/s and
% $\mathcal{V}_{\sigma^{\mathrm{y}}}(\omega,\alpha) = 606.4$ Km/s,
% corresponding to $\gamma_{\mathrm{x}}(\omega,\alpha) = 116.3^{\circ}$
% $\gamma_{\mathrm{y}}(\omega,\alpha) = 110.1^{\circ}$ and both for
% $\alpha=150^{\circ}$, where the plots are not presented.

@@

We reported the results of an \emph{ab initio} calculations for the spin
velocity injection (SVI) due to the one-photon absorption of the linearly
polarized light in the Up and Alt 2D 50\% hydrogenated graphene structures.
Different possible arrangements of the of the spin injection have been
considered: we made the calculations for the cases when the spin is polarized
in $z$ direction or when the velocity is directed along $x$ or $y$. To the best
of our knowledge, this effect  has not been previously reported in these  2D
partially hydrogenated structures. We have shown that the  SVI demonstrates an
anisotropic behavior, which is very sensitive to the symmetry of the structures
of interest and their band structure. % Do you think that the band structure 
% peculiarities are important? 
We have found that the Up structure shows the strongest response
for the spin directed along $z$, resulting in the velocity
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha) = 668.0$\,Km/s for the
incoming photon energy of 0.084\,eV. Also, the Alt structure has the strongest
response when the spin moves along $y$ direction, resulting in
$\mathcal{V}^{\mathrm{y}} (\omega,\alpha) = 905.6$\,Km/s for the incoming
photon energy of 0.720\,eV. The speed values obtained here are of the same
order of magnitude as those of Ref.~\onlinecite{najmaiePRB03} in unbiased
semiconductor quantum well structures, while they are of order of magnitude
higher compared to 3D bulk materials. Considering the fact that the spin relaxation time in pure and
doped graphene ranges from nanoseconds to milliseconds,
\cite{wojtaszekPRB13,ertlerPRB09}, and in view of the
high spin velocity transport that we obtained for the both
structures, this time is sufficiently long enough to have the SVI effect observed experimentally. Therefore, the Up and the Alt graphene structures
considered here are  excellent
candidates for the development of spintronics devices that require pure spin
current (PSC).

\section{Acknowledgment}
This work has been supported by \emph{Consejo Nacional de Ciencia y
Tecnolog\'ia} (CONACyT), M\'exico, Grant No. 153930.
R.Z.P. thanks CONACyT for scholarship support.

\bibliography{article.bib}

@@

>>>>

%\documentclass[prb,11pt,tightenlines,twocolumn,aps]{revtex4-1}
%\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
\documentclass[preprint,prb,showpacs]{revtex4}
\usepackage{amsmath}    % need for subequations 
\usepackage{graphicx}   % need for figures 
\usepackage{verbatim}   % useful for program listings 
\usepackage{color}      % use if color is used in text 
\usepackage{subfigure}  % use for side-by-side figures 
\usepackage[pdffitwindow,colorlinks,citecolor={red},linkcolor={blue}]{hyperref}
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage{soul}
\def\tama{10cm}
\raggedbottom           % don't add extra vertical space

@@

\title{Pure Spin Current Injection in Hydrogenated Graphene Structures}
\author{Reinaldo Zapata-Pe\~na}
\affiliation{Centro de Investigaciones en \'Optica, Le\'on,
Guanajuato 37150, M\'exico}
\author{Bernardo S. Mendoza}
\email[E-mail: ]{bms@cio.mx}
\affiliation{Centro de Investigaciones en \'Optica, Le\'on,
Guanajuato 37150, M\'exico}
\author{Anatoli I. Shkrebtii}
\affiliation{University of Ontario, Institute of Technology,
Oshawa, ON, L1H 7L7, Canada}

@@

\pacs{75.76+j,85.75.-d,78.67.Wj,78.90.+t}

@@

Spintronics is an emerging research field of electronics in which the
manipulation and transport of the electron spin in a solid state materials  is
central, adding a new degree of freedom to conventional charge
manipulation.\cite{wolfSC04,fabianAPS07} At present, there is an increasing
interest in attaining the same level of control over the transport of spin at
micro- or nano-scales, as it has been done for the flow of charge in typical
3D-bulk based electronic devices.\cite{awschalomNP2007} Several semiconductor
spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC). One of
the difficulties in creating measurable  spin current and
development of PSC based semiconductor devices is the fact that the spin
relaxation time in conventional semiconducting materials cloud be too short to enable the
spin transport, and may result in a non-observable spin
current.\cite{murakamiSc03} For PSC there is no net motion of charge; spin-up
electrons move in a given direction, while spin-down electrons travel in the
opposite one. This effect can be due to one-photon absorption of linearly
polarized light by a semiconductor, with filled valence bands and empty
conduction bands, illuminated by light with photon energy larger than the
energy gap. This phenomenon can be due to spin injection,\cite{malPRB03} Hall
Effects,\cite{sinovaPRB04} interference of two optical beams,\cite{bhatPRL00,
najmaiePRB03} or one photon absorption of linearly polarized
light\cite{bhatPRL05}. The last effect has been observed in gallium arsenide
(GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

@@

In the top panel of Fig. \ref{fig:up-vsz-w1}, we plot
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ {\it vs.}
0.080\,eV$\leq\hbar\omega\leq$0.096\,eV (similar energy range for the Up
structure shown in the left panel of Fig.~\ref{fig:vab-str-comp}) and
$0^\circ\leq\alpha\leq 180^\circ$. We see a broad peak that reaches the maximum
of $\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 739.7$\,Km/s at
$\alpha=35^{\circ}$ and $\hbar\omega= 0.084$\,eV.  The variation of
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ as a function of $\alpha$,
which comes from the interplay of the $\boldsymbol{\mu}$ tensor components as
multiplied by the  trigonometric functions of Eq.~\eqref{eq:vab-aw}, gives a
sizable set of values between 739.7\,Km/s and 165.4\,Km/s, for
0.084\,eV$\leq\hbar\omega\leq$0.090\,eV. In the bottom panel, we show
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ (left scale,
black solid line) {\it vs.} $\alpha$, at $\hbar\omega= 0.084$\,eV, thus following the ridge in the 3D
plot of the top panel. Also, we plot the corresponding velocity angle
$\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)$ (right scale, red
short-dashed line), where it
is very interesting to see that $\gamma_{\sigma^z}(\omega,\alpha)$ is centered
at 64.55$^\circ$ with a rather small deviation of only $\pm 0.03^\circ$, for
the whole range of $\alpha$. This result means that for $\hbar\omega=0.084$ eV
and for all values of $\alpha$, the electrons, with the chosen spin pointing
along $z$, will move at the angle of $\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)
\sim 64.5^{\circ}$ with respect to the $x$ direction, with the range of  high
speeds $\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ shown in the figure.
Also, from Eq. \eqref{eq:gamma-par} we find that
$\gamma^\parallel_{\sigma^\mathrm{z}} (\omega,\alpha)=\alpha = 64.56^\circ$,
with $\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha) = 631.1$\,Km/s 
(as indicated by the green dot-dashed arrow), and that from Eq. \eqref{eq:gamma-perp},
$\gamma^\perp_{\sigma^\mathrm{z}}(\omega,\alpha)=\alpha-90^\circ=64.50^\circ$,
gives $\alpha=154.50^\circ$, with
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 191.5$ Km/s 
(as indicated by the blue long-dashed 
arrow). 
Thus, at $\hbar\omega=0.084$ eV, an incident field, polarized at
$\alpha \sim 65.5^\circ$ or $\sim 154.5^\circ$, injects electrons with their
spin polarized along $z$, which move parallel or perpendicular to the incident
electric field,  with a speed of 631.14\,Km/s or 191.5\,Km/s, respectively.

Now, we analyze the results for the second energy range of the Up structure
shown in Fig.~\ref{fig:vab-str-comp}. In the top panel
of Fig. \ref{fig:up-vsz-w2}, we plot $\mathcal{V}_{\sigma^{\mathrm{z}}}
(\omega,\alpha)$ {\it vs.} 1.950\,eV$\leq\hbar\omega\leq$1.960\,eV and
$0^\circ\leq\alpha\leq 180^\circ$. We see two broad peaks that maximize at
$\alpha=35^{\circ}$ and $\hbar\omega= 1.954$\,eV, with a value of
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 193.5$\,Km/s, and at
$\alpha=35^{\circ}$ and $\hbar\omega= 1.957$\,eV, with a value of
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 170.6$\,Km/s. We only
analyze the highest  maximum in the bottom panel, where we  show
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ (left scale,
black solid line) {\it vs.} $\alpha$, at $\hbar\omega= 1.954$\,eV, thus following the highest ridge
shown in the 3D plot of the top panel. Also, we plot the corresponding velocity
angle $\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)$ (right scale, red short-dashed  line),
where in this case we see that the values of $\gamma_{\sigma^z}(\omega,\alpha)$
have more dispersion, as a function of $\alpha$, than for the lower energy
range shown in the bottom panel of Fig.~\ref{fig:up-vsz-w1}. However,
$\gamma_{\sigma^z}(\omega,\alpha)\sim 77.8^\circ$ is nearly constant from
$\alpha=0^\circ$ up to $\alpha\sim 85^\circ$. In this case, we find that
$\gamma^\parallel_{\sigma^\mathrm{z}}(\omega,\alpha)=\alpha=78.0^\circ$, with
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 115.0$\,Km/s 
(as indicated by the green dot-dashed arrow), and that from Eq. \eqref{eq:gamma-perp},
$\gamma^\perp_{\sigma^\mathrm{z}}(\omega,\alpha)=\alpha-90^\circ=167.8^\circ$,
gives $\alpha=77.8^\circ$, with
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 65.6$\,Km/s 
(as indicated by the blue long-dashed  arrow). 
Thus, through the correct choice of $\hbar\omega$ and $\alpha$ we could
inject electrons, in this case with their spin polarized along $z$, which move
parallel or perpendicular to the incident electric field, with sizable speeds.

@@

We proceed to analyze the Alt structure, just as we did with the Up structure,
but in this case, we only choose the lower energy range shown in the left
central panel of Fig.~\ref{fig:vab-str-comp}. In the top panel of Fig. 
\ref{fig:alt-vsz}, we plot $\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$
{\it vs.} 0.715\,eV$\leq\hbar\omega\leq$0.725\,eV and $0^\circ\leq\alpha\leq
180^\circ$. We see a broad peak that maximizes at $\alpha=150^{\circ}$ and
$\hbar\omega= 0.720$\,eV, with a value of
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 644.9$\,Km/s. In the bottom
panel, we  show $\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ (left scale, black solid line) {\it vs.}
$\alpha$, at $\hbar\omega= 0.720$\,eV, thus following
the highest ridge shown in the 3D plot of the top panel. Also, we plot the
corresponding velocity angle $\gamma_{\sigma^\mathrm{z}} (\omega,\alpha)$
(right scale, red short-dashed  line), where now we see that
$\gamma_{\sigma^z}(\omega,\alpha)$ is centered at $109.2^{\circ}$ having
variations of $\pm 1.0^{\circ}$ for $0^{\circ} \leq
\alpha \leq 180^{\circ}$. In this case, we find that
$\gamma^\parallel_{\sigma^\mathrm{z}} (\omega,\alpha) =
\alpha = 108.8^\circ$, with $\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)
= 450.05$\,Km/s (as indicated by the  green dot-dashed arrow), 
and that from Eq. \eqref{eq:gamma-perp},
$\gamma^\perp_{\sigma^\mathrm{z}}(\omega,\alpha)=\alpha-90^\circ=110.0^\circ$,
gives $\alpha=20.0^\circ$, with $\mathcal{V}_{\sigma^{\mathrm{z}}}
(\omega,\alpha) = 60.84$\,Km/s 
(as indicated by the blue long-dashed  arrow). Thus, as for the Up structure,
we could inject electrons with a fixed spin, which moves parallel or
perpendicular to the incident electric field.

@@

We reported the results of an \emph{ab initio} calculations for the spin
velocity injection (SVI) due to the one-photon absorption of the linearly
polarized light in the Up and Alt 2D 50\% hydrogenated graphene structures.
Different possible arrangements of the of the spin injection have been
considered: we made the calculations for the cases when the spin is polarized
in $z$ direction or when the velocity is directed along $x$ or $y$. 
To the best  
of our knowledge, this effect  has not been previously reported in these  2D  
partially hydrogenated structures.  
We have shown that the  SVI demonstrates an
anisotropic behavior, which is very sensitive to the symmetry of the structures
of interest.
We have found that the Up structure shows the strongest response
for the spin directed along $z$, resulting in the velocity
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha) = 668.0$\,Km/s for the
incoming photon energy of 0.084\,eV. Also, the Alt structure has the strongest
response when the spin moves along $y$ direction, resulting in
$\mathcal{V}^{\mathrm{y}} (\omega,\alpha) = 905.6$\,Km/s for the incoming
photon energy of 0.720\,eV. The speed values obtained here are of the same
order of magnitude as those of Ref.~\onlinecite{najmaiePRB03} in unbiased
semiconductor quantum well structures, while they are of order of magnitude
higher compared to 3D bulk materials. 
Considering the fact that the spin relaxation time in pure and
doped graphene ranges from nanoseconds to milliseconds,
\cite{wojtaszekPRB13,ertlerPRB09}, and in view of the
high spin velocity transport that we obtained for  both
structures, this time is sufficiently long enough to have the SVI
effect observed experimentally. 
Therefore, the Up and the Alt graphene structures
considered here are  excellent
candidates for the development of spintronics devices that require pure spin
current (PSC).

\section{Acknowledgment}
This work has been supported by \emph{Consejo Nacional de Ciencia y
Tecnolog\'ia} (CONACyT), M\'exico, Grant No. 153930.
R.Z.P. thanks CONACyT for scholarship support.
A.I.S thanks to Centro de Investigaciones en Optica (CIO) for the               
hospitality during his sabbatical research leave.

\bibliography{manuscript.bib}

@@

cover letter

We present a theoretical study of spin-velocity injection (SVI) of a pure spin 
current (PSC) induced by  a linearly polarized light that impinges normally on 
the surface of  two 50% hydrogenated noncentrosymmetric two-dimensional (2D) 
graphene structures. The first structure, hydrogenated at only one side, 
labeled Up, also known as graphone, and the second, labelled Alt, is 
25% hydrogenated at both sides. The hydrogenation opens an energy gap in both 
structures. We analyze two possibilities: in the first, the spin is fixed 
along a chosen direction, and the resulting SVI is calculated; 
in the second, we choose the SVI direction along 
the surface plane, and calculate the resulting 
spin orientation. 
The speed values that we obtain  are of order of magnitude larger than those of bulk semiconductors, 
such as CdSe and GaAs, thus making the hydrogenated graphene structures 
excellent candidates for spintronics applications. 
To the best  
of our knowledge, this effect  has not been previously reported in these  2D  
partially hydrogenated structures.  
We have published several related works in Physical Review B with very
good visibility, for which we know that the present manuscript is
suitable for PRB as well.
